---
title: "How to Wrap Julia Packages in R: A Step-by-Step Guide"
author: "Your Name"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Wrap Julia Packages in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette provides a comprehensive guide to wrapping Julia packages in R, using the `rjfiglet` package as a case study. By the end, you'll understand how to:

1. Set up R-Julia communication with `JuliaCall`
2. Wrap Julia functions in R functions
3. Handle data type conversions
4. Manage dependencies and initialization
5. Create a complete R package around Julia functionality

## Prerequisites

Before starting, ensure you have:

- R (>= 4.0.0)
- Julia (>= 1.6) installed and in PATH
- Basic knowledge of R package development
- Familiarity with both R and Julia syntax

## Step 1: Project Setup

### Create R Package Structure

Start by creating a new R package. You can use `devtools` or `usethis`:

```r
usethis::create_package("myjuliawrapper")
```

Or manually create the basic structure:

```
myjuliawrapper/
├── DESCRIPTION
├── NAMESPACE
├── R/
└── man/
```

### Configure DESCRIPTION

Add `JuliaCall` to the `Imports` field and specify system requirements:

```yaml
Package: myjuliawrapper
Imports:
    JuliaCall
SystemRequirements: Julia (>= 1.6)
```

## Step 2: Initialize Julia Connection

### Package Loading Hook

Create `R/zzz.R` (or any R file) with `.onLoad` to initialize Julia:

```{r}
# Example from rjfiglet
.onLoad <- function(libname, pkgname) {
  # Setup Julia
  JuliaCall::julia_setup()
  
  # Check if target Julia package is installed
  pkg_name <- "TargetPackage"
  if (!JuliaCall::julia_eval(sprintf('Base.find_package("%s") !== nothing', pkg_name))) {
    message(sprintf("Installing %s.jl via Pkg...", pkg_name))
    JuliaCall::julia_eval(sprintf('using Pkg; Pkg.add("%s")', pkg_name))
  }
  
  # Load the package
  JuliaCall::julia_eval(sprintf('using %s', pkg_name))
}
```

**Key points:**

- `julia_setup()` initializes the Julia session
- Check for package existence before attempting to install
- Install automatically if missing (optional, can be user-controlled)
- Load the package for immediate use

## Step 3: Wrap Julia Functions

### Basic Function Wrapping

Consider a Julia function `TargetPackage.foo(x, y)`. Here's how to wrap it:

```{r}
#' Wrapped version of TargetPackage.foo
#'
#' @param x,y Numeric inputs
#' @return Result from Julia
#' @export
foo <- function(x, y) {
  # Convert R inputs to Julia-compatible form
  julia_x <- JuliaCall::julia_eval(as.character(x))
  julia_y <- JuliaCall::julia_eval(as.character(y))
  
  # Call Julia function
  result <- JuliaCall::julia_eval(sprintf('TargetPackage.foo(%s, %s)', 
                                          julia_x, julia_y))
  
  # Convert result back to R
  return(result)
}
```

### Advanced Wrapping with Multiple Parameters

For functions with many parameters, use named arguments and construct the Julia call dynamically:

```{r}
#| eval: false

#' Advanced wrapper with multiple parameters
#'
#' @param text Character string
#' @param font Font name
#' @return ASCII art lines
#' @export
advanced_wrapper <- function(text, font = "standard") {
  
  # Start building Julia command
  cmd <- sprintf('TargetPackage.func("%s", font="%s")', text, font)
      
  # Execute and return
  result <- JuliaCall::julia_eval(cmd)
  return(result)
}
```

## Step 4: Data Type Conversion

### Conversion Table

| R Type | Julia Type | Conversion Method |
|--------|------------|-------------------|
| `character` | `String` | `JuliaCall::julia_eval(paste0('"', x, '"'))` |
| `numeric` | `Float64` | `JuliaCall::julia_eval(as.character(x))` |
| `integer` | `Int64` | `JuliaCall::julia_eval(as.character(x))` |
| `logical` | `Bool` | `JuliaCall::julia_eval(ifelse(x, "true", "false"))` |
| `vector` | `Array` | `JuliaCall::julia_call("collect", x)` |
| `matrix` | `Matrix` | `JuliaCall::julia_call("Matrix", x)` |
| `list` | `Dict` | More complex; may need manual conversion |

### Example Conversions

```{r}
# Character to Julia String
r_text <- "Hello"
julia_text <- JuliaCall::julia_eval(paste0('"', r_text, '"'))

# Numeric vector to Julia Array
r_vec <- c(1, 2, 3)
julia_vec <- JuliaCall::julia_call("collect", r_vec)

# Get result back to R
  julia_result <- JuliaCall::julia_eval("[1, 2, 3] .* 2")
  r_result <- julia_result  # c(2, 4, 6)
```

## Step 5: Error Handling and Validation

### Input Validation

```{r}
safe_wrapper <- function(x, y) {
  # Validate inputs
  if (!is.numeric(x) || !is.numeric(y)) {
    stop("Both x and y must be numeric")
  }
  if (length(x) != 1 || length(y) != 1) {
    stop("x and y must be scalars")
  }
  
  # Check Julia is ready
  if (!yulab.utils:::has_bin('julia')) {
    stop("Julia not initialized. Call julia_setup() first.")
  }
  
  # Try the Julia call with error handling
  tryCatch({
    result <- JuliaCall::julia_eval(sprintf('TargetPackage.foo(%f, %f)', x, y))
    result
  }, error = function(e) {
    stop(sprintf("Julia error: %s", e$message))
  })
}
```

### Julia Package Availability

```{r}
check_julia_package <- function(pkg_name) {
  tryCatch({
    JuliaCall::julia_eval(sprintf('using %s', pkg_name))
    TRUE
  }, error = function(e) {
    FALSE
  })
}
```

## Step 6: Performance Considerations

### Caching Julia Results

For expensive Julia computations, consider caching:

```{r eval=FALSE}
# Simple caching with memoise
library(memoise)

cached_julia_computation <- memoise(function(x) {
  JuliaCall::julia_eval(sprintf("expensive_computation(%f)", x))
})
```

### Batch Operations

Instead of calling Julia repeatedly in a loop, pass vectorized operations:

```{r}
# 首先定义process函数
JuliaCall::julia_eval("function process(x)
  return x^2
end")

# Inefficient
results <- sapply(1:10, function(i) {
  JuliaCall::julia_eval(sprintf("process(%d)", i))
})

# More efficient
results <- JuliaCall::julia_eval("[process(i) for i in 1:10]")
results
```

## Step 7: Testing Your Wrapper

### Test Structure

Create tests in `tests/testthat/`:

```r
# tests/testthat/test-wrapper.R
test_that("basic wrapper works", {
  skip_if_not(yulab.utils:::has_bin('julia'), "Julia not available")
  
  result <- foo(2, 3)
  expect_type(result, "double")
  expect_length(result, 1)
})

test_that("error handling works", {
  expect_error(foo("not numeric", 3))
})
```

### Skip Tests Without Julia

Use `skip_if_not()` to avoid failures when Julia isn't available:

```r
skip_if_not(yulab.utils:::has_bin('julia'), "Julia not installed or configured")
```

## Step 8: Documenting Your Package

### Roxygen Documentation

Document each wrapped function with details about the underlying Julia function:

```r
#' Compute something using TargetPackage.foo
#' 
#' This function wraps `TargetPackage.foo` from Julia. See the
#' [Julia documentation](https://targetpackage.readthedocs.io/) for details.
#'
#' @param x,y Numeric inputs passed to `TargetPackage.foo`
#' @return The result from Julia computation
#' @examples
#' if (interactive() && JuliaCall::julia_setup_ok()) {
#'   foo(2, 3)
#' }
#' @export
foo <- function(x, y) {
  # ... implementation
}
```

### Vignettes and README

- Explain why users should use your wrapper
- Provide comparison with pure R alternatives (if any)
- Include benchmarks showing performance benefits
- Document any limitations or caveats

## Case Study: rjfiglet Implementation

Let's examine how `rjfiglet` implements these principles:

### 1. Initialization

```r
# In R/rjfiglet.R
.onLoad <- function(libname, pkgname) {
  JuliaCall::julia_setup()
  # Auto-install FIGlet.jl if missing
  # ...
}
```

### 2. Main Wrapper Function

```r
# In R/figlet.R
figlet <- function(text, font = "standard") {
  # Ensure Julia is set up and FIGlet.jl loaded (handled by .onLoad)
  # Use JuliaCall::julia_eval with proper string escaping
  # Build the Julia expression
  expr <- sprintf('FIGlet.render("%s", "%s")', text, font)
  invisible(JuliaCall::julia_eval(expr))
}
```

### 3. Utility Functions

```r
# In R/utils.R
check_julia_setup <- function() {
  if (!yulab.utils:::has_bin('julia')) {
    stop("Julia is not set up...")
  }
  # ...
}
```

## Common Pitfalls and Solutions

### 1. Julia Not Found

**Problem**: `julia_setup()` fails because Julia isn't in PATH.

**Solution**:
- Check `JuliaCall::find_julia()`
- Set `JULIA_HOME` environment variable
- Provide installation instructions in README

### 2. Package Installation Failures

**Problem**: Auto-installation of Julia packages fails due to network or permissions.

**Solution**:
- Make installation optional with informative messages
- Provide manual installation instructions
- Use `tryCatch` with fallbacks

### 3. Memory Leaks

**Problem**: Large data transfers between R and Julia cause memory issues.

**Solution**:
- Transfer only necessary data
- Clear intermediate Julia objects when possible
- Use `JuliaCall::julia_eval("GC.gc()")` to trigger garbage collection

### 4. Type Conversion Errors

**Problem**: Complex R types don't convert correctly to Julia.

**Solution**:
- Convert to simpler types first (e.g., matrix → vector)
- Write custom conversion functions
- Accept only supported types in your wrapper

## Advanced Topics

### Custom Julia Modules

For complex wrappers, create a custom Julia module:

```julia
# src/mywrapper.jl
module MyWrapper

using TargetPackage

export wrapped_foo

function wrapped_foo(x, y)
    # Additional processing
    result = TargetPackage.foo(x, y)
    # Post-processing
    return result
end

end # module
```

Load it in R:

```r
JuliaCall::julia_source("src/mywrapper.jl")
```

### Parallel Execution

Julia can run computations in parallel. Expose this capability:

```r
parallel_computation <- function(inputs) {
  JuliaCall::julia_eval("using Distributed")
  JuliaCall::julia_eval("addprocs(4)")  # Add 4 processes
  
  # Distribute computation
  result <- JuliaCall::julia_eval(sprintf(
    "@distributed (+) for x in %s
        expensive_computation(x)
    end", 
    paste(inputs, collapse = ", ")
  ))
  
  return(JuliaCall::julia_get(result))
}
```

## Conclusion

Wrapping Julia packages in R provides access to high-performance numerical computing while maintaining R's ease of use and extensive ecosystem. The key steps are:

1. **Setup**: Initialize Julia and manage dependencies
2. **Wrapping**: Convert R functions to call Julia with proper type handling
3. **Validation**: Add error checking and user-friendly messages
4. **Documentation**: Explain both R interface and underlying Julia functionality
5. **Testing**: Ensure reliability across different environments

The `rjfiglet` package demonstrates these principles for a relatively simple package. For more complex Julia packages, you may need additional layers of abstraction, but the core concepts remain the same.

## Further Resources

- [JuliaCall documentation](https://cran.r-project.org/package=JuliaCall)
- [Julia documentation](https://docs.julialang.org/)
- [R Packages book](https://r-pkgs.org/)
- [Julia and R: Best of Both Worlds](https://www.juliabloggers.com/julia-and-r-best-of-both-worlds/)

## References

1. JuliaCall: Seamless Integration Between R and Julia (2019). Journal of Statistical Software.
2. Figlet.jl documentation: https://github.com/JuliaString/Figlet.jl
3. R package development guidelines: https://r-pkgs.org/

---

*This vignette was generated as part of the `rjfiglet` package. The package serves both as a useful tool and an educational example of R-Julia integration.*
